---
title: "Variogram"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

.wd <- '~/projects/tutorials'
source(file.path(.wd,'startup.r'))
source(file.path(.wd,'funs/themes.r'))

theme_set(theme_eda)

```

## Random walk

```{r}

#https://stackoverflow.com/questions/21991130/simulating-a-random-walk

n <- 1000

dat <- tibble(t=1:n,
  y=cumsum(sample(c(-1, 1), n, TRUE)))

ggplot(dat,aes(x=t,y=y)) + geom_line()

```

## Semi-variance

Compute semi-variance at lag 1.

See formula for empirical variogram: http://wiki.gis.com/wiki/index.php/Variogram

Since each observation was taken at a lag of 1, we take all observations.

```{r}

z <- combn(dat$y,2,simplify=TRUE) %>% t #Form every possible pair of observations

sum((z[,1] - z[,2])^2)/length(z) #Calculate semi-variance

```

Now calculate semi-variance for items at lag 2.

Take every other observation.

```{r}

y <- dat$y[seq(1,n,by=2)] #Take every other observation.

z <- combn(y,2,simplify=TRUE) %>% t #Form every possible pair of observations

sum((z[,1] - z[,2])^2)/length(z) #Calculate semi-variance

```

But, notice we could have taken two sets at lag 2. One set with observations (1,3,5,...) and another set with observations (2,4,6,...).

In general, as lag increases, we get more possible sets, but the number of observations in each set decreases.

## Forming sets of observations at lag = h

Assume n = 10 observations, each taken at lag of 1 (say one per minute from 1 to 10 minutes).

At lag of 1, we only have one set of size 10

At lag of 2, we have two sets, each of size 5

```{r}
seq(1,10,by=2)
seq(2,10,by=2)

```
At lag 3 we have three sets, with sizes 4, 3, 3

```{r}
seq(1,10,by=3)
seq(2,10,by=3)
seq(3,10,by=3)
```

At lag 4 we have four sets, with sizes 3,3,2,2

```{r}
seq(1,10,by=4)
seq(2,10,by=4)
seq(3,10,by=4)
seq(4,10,by=4)
```
So the number of sets increases but set size decreases until eventually there are 10 sets of size 1

```{r}
lag=10
for(i in 1:lag) {
  print(seq(i,10,by=lag))
}
  
```
So we only do `max lag/2` ? So in this case max lag would be 5?

```{r}

n <- 10
lags <- 1:5

svh <- tibble()

for(lag in lags) {
  
  for(i in 1:lag) {
    
    y <- dat[seq(i,n,by=lag),]$y
    
    z <- combn(y,2,simplify=TRUE) %>% t

    sv <- sum((z[,1] - z[,2])^2)/length(z)
    
    svh <- bind_rows(svh,tibble(lag=lag,sv=sv))
  }
}

```

Plot semivariance by lag (points), and a line for mean semivariance.

```{r}

ggplot(svh,aes(x=lag,y=sv)) + 
  geom_point() +
  geom_line(dat=svh %>% group_by(lag) %>% summarize(sv=mean(sv)))

```

Put it all together with larger number of points.

```{r}

n <- 100
lags <- 1:(n/2)
svh <- tibble()

dat <- tibble(t=1:n,
  y=cumsum(sample(c(-1, 1), n, TRUE)))

tic()
for(lag in lags) {
  #lag=2
  for(i in 1:lag) {
    #i=1
    y <- dat[seq(i,n,by=lag),]$y
    
    z <- combn(y,2,simplify=TRUE) %>% t

    sv <- sum((z[,1] - z[,2])^2)/length(z)
    
    svh <- bind_rows(svh,tibble(lag=lag,sv=sv))
  }
}
toc()

ggplot(svh,aes(x=lag,y=sv)) + 
  geom_point() +
  geom_line(dat=svh %>% group_by(lag) %>% summarize(sv=mean(sv)))

```
